Control Aware Radio Resource Allocation in Low Latency Wireless Control Systems

### 2019-02-18 Low_latency_control/revised/introduction.tex
Add files via upload

In the context of wireless control systems, there have been a range of works that incorporates control system information in the networking and communication policies.
%
The mechanisms usually examined are either static or dynamic. Typical examples of the former type are periodically protocols where the wireless devices transmit in a predefined repeating order, e.g., round-robin. Control system stability under such protocols can be analyzed -- see, e.g.,~\cite{Hespanha_survey, Schenato_foundations, Donkers_switched, Branicky_stability}. Periodic sequences leading to stability~\cite{Hristu_shared_feedback}, controllability and observability~\cite{Hristu_communication_control}, or optimizing control objectives~\cite{LeNy_resource_LQR,Meier_measurement_control, Scheduling_control_combinatorics} have been proposed.
%
Dynamic schedulers do not rely on a predefined sequence but decide access to the communication medium dynamically at each step. %, for example by dynamically assigning priorities to the competing tasks.
Initial approaches abstract control performance requirements in the time/frequency domain, e.g., how often a task needs resource access, employing algorithms from real-time scheduling theory~\cite{Branicky_RM, Liu_Real_time_systems}. More recent scheduling approaches often depend on the current control system states, i.e., informally the subsystem with the largest state discrepancy is scheduled to communicate -- see, e.g.,~\cite{Donkers_switched,  Cervin_event_scheduling, mamduhi2014event,shi2011optimal,han2017optimal}.
%
Alternatively scheduling can take into account current wireless channel conditions opportunistically to meet target control system reliability requirements~\cite{GatsisEtal15}. \red{None of these approaches, however, are explicitly designed to achieve strong performance it low-latency scenarios, which are critical in modern industrial control and factory settings \cite{brahmi2015deployment,yilmaz2015analysis}.}

>>>>

In the context of wireless control systems, there have been a range of works that incorporates control system information in the networking and communication policies.
%
The mechanisms usually examined are either static or dynamic. Typical examples of the former type are periodically protocols where the wireless devices transmit in a predefined repeating order, e.g., round-robin. Control system stability under such protocols can be analyzed -- see, e.g.,~\cite{Hespanha_survey, Schenato_foundations, Donkers_switched, Branicky_stability}. Periodic sequences leading to stability~\cite{Hristu_shared_feedback}, controllability and observability~\cite{Hristu_communication_control}, or optimizing control objectives~\cite{LeNy_resource_LQR,Meier_measurement_control, Scheduling_control_combinatorics} have been proposed.
%
Dynamic schedulers do not rely on a predefined sequence but decide access to the communication medium dynamically at each step. %, for example by dynamically assigning priorities to the competing tasks.
Initial approaches abstract control performance requirements in the time/frequency domain, e.g., how often a task needs resource access, employing algorithms from real-time scheduling theory~\cite{Branicky_RM, Liu_Real_time_systems}. More recent scheduling approaches often depend on the current control system states, i.e., informally the subsystem with the largest state discrepancy is scheduled to communicate -- see, e.g.,~\cite{Donkers_switched,  Cervin_event_scheduling, mamduhi2014event,shi2011optimal,han2017optimal}.
%
Alternatively scheduling can take into account current wireless channel conditions opportunistically to meet target control system reliability requirements~\cite{GatsisEtal15}. \red{None of these approaches, however, are explicitly designed to achieve strong performance in large scale low-latency systems, which are critical in modern industrial control and factory settings \cite{brahmi2015deployment,yilmaz2015analysis}.}

### 2019-02-18 Low_latency_control/revised/scheduling_alg.tex
Add files via upload

>>>>

\red{
 \begin{remark}\label{remark_estimation}\normalfont
In practical systems, the channel state information $\bbH_k$ is often obtained with some estimation errors, which may impact the channel aware scheduling approach taken here. Observer, however, that the computation of adaptive PDR targets in \eqref{eq_pdr_constraint} does not depend upon channel state information. Thus, the primary component of the CALLS method---namely Steps 3-6 in Algorithm \ref{alg_calls}--- is unaffected by inaccurate channel estimation. Likewise, the assignment method-based scheduling in Step 8 does also not directly depend upon channel information---see the formulation of the assignment problem in \eqref{eq_assignment}. The only component that may be negatively impacted by estimation errors is Step 7, i.e.  the maximum MCS selection in \eqref{eq_mcs_select}. Here, estimation errors may result in the selection of an MCS that cannot meet the target PDR targets.  Such an effect can be mitigated by selecting a more conservative MCS selection to account for channel estimation errors---e.g. transmitting at an MCS $\mu_{i,k}(\bbvarsigma) - 1$ rather than $\mu_{i,k}(\bbvarsigma)$. Such a conservative scheme would tradeoff latency to the benefit of reliability or robustness.
 \end{remark}
 }

### 2019-02-18 Low_latency_control/revised/simulations.tex
Add files via upload

We perform an initial set of simulations on the well-studied problem of controlling a series of inverted pendulums on a horizontal cart. While conceptually simple, the highly unstable dynamics of the inverted pendulum make it a representative example of control system that requires fast control cycles, and subsequently low-latency communications when being controlled over a wireless medium. Consider a series of $m$ identical inverted pendulums, as pictured in Figure \ref{fig_inverted_pendulum}. Each pendulum of length $L$ is attached at one end to a cart that can move along a single, horizontal axis. The position of the pendulum changes by the effects of gravity and the force applied to the linear cart. For our experiments, we use the modeling of the inverted pendulum as provided by Quanser \cite{quanser}. The state is $p=4$ dimensional vector that maintains the position and velocity of the cart along the horizontal axis, and the angular position and velocity of the pendulum, i.e. $\bbx_{i,k} := [x_{i,k}, \dot{x}_{i,k}, \theta_{i,k}, \dot{\theta}_{i,k}]$. The system input $u_{i,k}$ reflects a horizontal force placed upon $i$th pendulum. By applying a zeroth order hold on the continuous dynamics with a state sampling rate of $0.01$ seconds and linearizing, we obtained the following discrete linear dynamic matrices of the pendulum system
%
\begin{align}\label{eq_control_orig}
\bbA_i =
\begin{bmatrix}
1 & 0 & 0 & 0 \\
0 & 2.055 & -0.722 & 4.828 \\
0 & 0.023 & 0.91 & 0.037 \\
0 & 0.677 & -0.453 & 2.055
\end{bmatrix},
\bbB_i =
\begin{bmatrix}
0.034 \\ 0.168 \\ 0.019 \\ 0.105
\end{bmatrix}.
\end{align}
%

@@

%%%%%%
\begin{figure}
\centering
\begin{tikzpicture}[scale=0.8]
\begin{axis}[
    ybar,
    enlargelimits=0.15,
    legend style={at={(0.5,0.85)},
      anchor=north,legend columns=-1},
    ylabel={\# devices supported},
    symbolic x coords={0.5 ms,1 ms,1.5 ms},
    xtick=data,
    nodes near coords,
    nodes near coords align={vertical},
    ]
\addplot coordinates {(0.5 ms,9) (1 ms,19) (1.5 ms,36)};
\addplot coordinates {(0.5 ms,12) (1 ms,28) (1.5 ms,55)};
\legend{Fixed PDR, CALLS}
\end{axis}
\end{tikzpicture}
\caption{Total number of inverted pendulum devices that can be controlled using Fixed-PDR and CALLS scheduling for various latency thresholds.}
\label{fig_ip_bar}
\end{figure}
%%%%%%

@@

In Figure \ref{fig_bbb_dist} we show the results of a representative simulation of the control of $m=50$ balancing board ball systems with a latency bound of $\tau_{\max} = 10^{-3}$ seconds. Observe that, in this system, even with a large number of users the CALLS method can keep all systems very close to the center of the board, while the fixed PDR scheduler loses a few of the balls due to the agnosticism of the scheduler.

 To dive deeper into the benefits provided by control aware scheduling, we present in Figure \ref{fig_bbb_psr} a histogram of the actual  packet delivery rates each of the devices achieved over the representative simulation. It is interesting to observe that, in the CALLS method, the achieved PDRs are closely concentrated, ranging from 0.3 to 0.44. On the other hand, using a fixed PDR scheduling scheme, the non-variable rates are too strict for the low-latency system to support, and without control-aware scheduling the achieved PDRs range wildly from close to 0 to close to 1. In this case, some devices are able to transmit almost every cycle while others are almost never able to successfully transmit their packets. This suggests that, by using control aware scheduling, we indirectly achieve a sense of fairness across users over the long term. Further note that the PDRs required to keep the balancing board ball stable, e.g. 0.4, are relatively small. This is due to the fact that the balancing board ball features relatively slow moving dynamics, making it easier to control with less frequent transmissions. This is comparison to the inverted pendulum system, in which the pendulums were kept stable with PDRs in the range 0.6-0.75.

%%%%
\begin{figure}
\centering
\includegraphics[width=.45\textwidth]{../images/bbb_dist.eps}
\caption{Average ball distance to center for $m=50$ devices using (top) CALLS and (bottom) fixed-PDR scheduling with $\tau_{\max} =1$ ms latency threshold. The proposed control aware scheme keeps all balancing balls close to center, while fixed-PDR scheduling cannot.}
\label{fig_bbb_dist}
\end{figure}
%%

%%%%
\begin{figure}
\centering
\includegraphics[width=.45\textwidth]{../images/bbb_psr.eps}
\caption{Histogram of achieved PDRs in $m=50$ balancing board systems (top) CALLS and (bottom) fixed-PDR scheduling with $\tau_{\max} =1$ ms latency threshold. The proposed control aware scheme achieves similar PDRs for all devices, while the control agnostic scheduling results in large variation in packet delivery.}
\label{fig_bbb_psr}
\end{figure}
%%

We present in Figure \ref{fig_bbb_bar} the final capacity results obtained over all the simulations for the balancing board ball system. Observe that proposed approach increases the number of supported devices by factor of 2 relative to the standard fixed PDR approach. The even greater improvement here relative to the inverted pendulum simulations can be attributed to the slower dynamics of the balancing board ball, which allows for even more gains using control-aware PDRs due to the lower PDR requirements of the system.

%%%%%%
\begin{figure}
\centering
\begin{tikzpicture}[scale=0.8]
\begin{axis}[
    ybar,
    enlargelimits=0.15,
    legend style={at={(0.5,0.85)},
      anchor=north,legend columns=-1},
    ylabel={\# devices supported},
    symbolic x coords={0.5 ms,1 ms},
    xtick=data,
    nodes near coords,
    nodes near coords align={vertical},
    ]
\addplot coordinates {(0.5 ms,9) (1 ms,27)};
\addplot coordinates {(0.5 ms,25) (1 ms,60)};
\legend{Fixed PDR, CALLS}
\end{axis}
\end{tikzpicture}
\caption{Total number of balancing ball board devices that can be controlled using Fixed-PDR and CALLS scheduling for various latency thresholds.}
\label{fig_bbb_bar}
\end{figure}
%%%%%%

>>>>

We perform an initial set of simulations on the well-studied problem of controlling a series of inverted pendulums on a horizontal cart. While conceptually simple, the highly unstable dynamics of the inverted pendulum make it a representative example of control system that requires fast control cycles, and subsequently low-latency communications when being controlled over a wireless medium. Consider a series of $m$ identical inverted pendulums, as pictured in Figure \ref{fig_inverted_pendulum}. Each pendulum of length $L$ is attached at one end to a cart that can move along a single, horizontal axis. The position of the pendulum changes by the effects of gravity and the force applied to the linear cart. For our experiments, we use the modeling of the inverted pendulum as provided by Quanser \cite{quanser}. The state is $p=4$ dimensional vector that maintains the position and velocity of the cart along the horizontal axis, and the angular position and velocity of the pendulum, i.e. $\bbx_{i,k} := [x_{i,k}, \dot{x}_{i,k}, \theta_{i,k}, \dot{\theta}_{i,k}]$. The system input $u_{i,k}$ reflects a horizontal force placed upon $i$th pendulum. By applying a zeroth order hold on the continuous dynamics with a state sampling rate of $0.01$ seconds and linearizing, we obtained the following discrete linear dynamic matrices of the pendulum system
%
\begin{align}\label{eq_control_orig}
\bbA_i =
\begin{bmatrix}
1 & 0.037 & 3.477 & 0.042 \\
0 & 2.055 & -0.722 & 4.828 \\
0 & 0.023 & 0.91 & 0.037 \\
0 & 0.677 & -0.453 & 2.055
\end{bmatrix},
\bbB_i =
\begin{bmatrix}
0.034 \\ 0.168 \\ 0.019 \\ 0.105
\end{bmatrix}.
\end{align}
%

@@

%%%%%%
\begin{figure}
\centering
\begin{tikzpicture}[scale=0.8]
\begin{axis}[
    ybar,
    enlargelimits=0.15,
    legend style={at={(0.5,0.85),size=\footnotesize},
      anchor=north,legend columns=-1},
    ylabel={\# devices supported},
    symbolic x coords={0.5 ms,1 ms,1.5 ms},
    xtick=data,
    nodes near coords,
    nodes near coords align={vertical},
    ]
\addplot[blue!40!black,fill=blue!20!white] coordinates {(0.5 ms,9) (1 ms,19) (1.5 ms,36)};
%\addplot[green!40!black,fill=green!20!white] coordinates {(0.5 ms,9) (1 ms,19) (1.5 ms,36)};
\addplot[red!40!black,fill=red!20!white] coordinates {(0.5 ms,12) (1 ms,28) (1.5 ms,55)};
\legend{Fixed PDR,  CALLS}
\end{axis}
\end{tikzpicture}
\caption{Total number of inverted pendulum devices that can be controlled using Fixed-PDR and CALLS scheduling for various latency thresholds.}
\label{fig_ip_bar}
\end{figure}
%%%%%%

@@

\red{
In the simulations performed with the balancing board system, in addition to making comparisons of the CALLS method to a fixed PDR low latency scheduling scheme, we perform additional comparisons to a standard control-aware scheduling approach---namely, the event-triggered scheduling approach \cite{Cervin_event_scheduling, mamduhi2014event}. In event triggered scheduling, we schedule devices only when its estimated control state goes above some threshold value. When such an event occurs, this device is scheduled with a fixed high reliability PDR using a low-latency assignment based scheduling method. This, in effect, combines the selective scheduling approach of CALLS with fixed high reliability PDR targets commonly used in URLLC.

In Figure \ref{fig_bbb_dist} we show the results of a representative simulation of the control of $m=50$ balancing board ball systems with a latency bound of $\tau_{\max} = 10^{-3}$ seconds. Observe that, in this system, even with a large number of users, both the event-triggered scheduling and the CALLS method can keep all systems very close to the center of the board, while the fixed PDR scheduler loses a few of the balls due to the agnosticism of the scheduler.

 To dive deeper into the benefits provided by control aware scheduling, we present in Figure \ref{fig_bbb_psr} a histogram of the actual  packet delivery rates each of the devices achieved over the representative simulation. It is interesting to observe that, in both the event triggered and CALLS method, the achieved PDRs are closely concentrated, ranging from 0.3 to 0.44. On the other hand, using a fixed PDR scheduling scheme, the non-variable rates are too strict for the low-latency system to support, and without control-aware scheduling the achieved PDRs range wildly from close to 0 to close to 1. In this case, some devices are able to transmit almost every cycle while others are almost never able to successfully transmit their packets. This suggests that, by using control aware scheduling, we indirectly achieve a sense of fairness across users over the long term. Further note that the PDRs required to keep the balancing board ball stable, e.g. 0.4, are relatively small. This is due to the fact that the balancing board ball features relatively slow moving dynamics, making it easier to control with less frequent transmissions. This is comparison to the inverted pendulum system, in which the pendulums were kept stable with PDRs in the range 0.6-0.75.
 }

%%%%
\begin{figure}
\centering
\includegraphics[width=.45\textwidth]{../images/bbb_dist.eps}
\caption{Average ball distance to center for $m=50$ devices using (top) CALLS, (middle) event-triggered, and (bottom) fixed-PDR scheduling with $\tau_{\max} =1$ ms latency threshold. The proposed control aware scheme keeps all balancing balls close to center, while fixed-PDR scheduling cannot.}
\label{fig_bbb_dist}
\end{figure}
%%

%%%%
\begin{figure}
\centering
\includegraphics[width=.45\textwidth]{../images/bbb_psr.eps}
\caption{Histogram of achieved PDRs in $m=50$ balancing board systems (top) fixed PDR, (middle) event-triggered, and (bottom) CALLS scheduling with $\tau_{\max} =1$ ms latency threshold. The proposed control aware scheme achieves similar PDRs for all devices, while the control agnostic scheduling results in large variation in packet delivery.}
\label{fig_bbb_psr}
\end{figure}
%%

\red{
We present in Figure \ref{fig_bbb_bar} the final capacity results obtained over all the simulations for the balancing board ball system. Observe that proposed approach increases the number of supported devices by factor of 2 relative to the standard fixed PDR approach. The even greater improvement here relative to the inverted pendulum simulations can be attributed to the slower dynamics of the balancing board ball, which allows for even more gains using control-aware PDRs due to the lower PDR requirements of the system. Likewise, the Lyapunov-based adaptive PDR requirements allow for even greater scalability than the more standard event-triggered approach, which can service only 17 and 50 users with 0.5 and 1 ms latency thresholds, respectively.}

%%%%%%
\begin{figure}
\centering
\begin{tikzpicture}[scale=0.8]
\begin{axis}[
    ybar,
    enlargelimits=0.15,
    legend style={at={(0.4,.85),font=\footnotesize},
      anchor=north,legend columns=-1},
    ylabel={\# devices supported},
    symbolic x coords={0.5 ms,1 ms},
    xtick=data,
    nodes near coords,
    nodes near coords align={vertical},
    ]
\addplot[blue!40!black,fill=blue!20!white] coordinates {(0.5 ms,9) (1 ms,27)};
\addplot[green!40!black,fill=green!20!white] coordinates {(0.5 ms,17) (1 ms,50)};
\addplot[red!40!black,fill=red!20!white] coordinates {(0.5 ms,25) (1 ms,60)};
\legend{Fixed PDR, Event-Trigged, CALLS}
\end{axis}
\end{tikzpicture}
\caption{Total number of balancing ball board devices that can be controlled using Fixed-PDR, Event-Triggered, and CALLS scheduling for various latency thresholds.}
\label{fig_bbb_bar}
\end{figure}
%%%%%%
